---
title: "R Notebook"
output: html_notebook
---

# p1: strictly continuous only

```{r}

p1 <- pca_sak_continuous_only_z_scored$ind$coord |> 
  as_tibble() |> 
  rename(x = Dim.1, y = Dim.2) |> 
  mutate(class = df_sak$class, name = df_sak$short_name) |> 
  ggplot(aes(x, y, colour = class |> as.factor(), label = class)) +
  geom_text(size = 2, family = "Fira Code") +
  theme_pokemon +
  labs(
    x = paste0(
      "PCA 1, ",
      pca_sak_continuous_only_z_scored$eig[1,2] |> round(digits = 1),
      "%"),
    y =  paste0(
      "PCA 2, ", 
      pca_sak_continuous_only_z_scored$eig[2,2] |> round(digits = 1), 
      "%"
      ),
    title = "PCA continuous only"
  ) +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )
```
余力あったらTODO: clean up ticks

# p2: all cont TSNE

```{r}
p2 <- tsne_sak_continuous_only_z_scored |> 
  as_tibble() |> 
  rename(x = V1, y = V2) |> 
  mutate(class = df_sak$class, name = df_sak$short_name) |> 
  ggplot(aes(x, y, colour = class |> as.factor(), label = class)) +
  geom_text(size = 2, family = "Fira Code") +
  labs(
    title = "t-SNE continuous only",
    x = expression(paste("t-SNE  ", italic(X))),
    y = expression(paste("t-SNE  ", italic(Y))),
  ) +
  theme_pokemon +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )
  
```


# p3: UMAP cont only

```{r}
p3 <- umap_sak_continuous_only_z_scored |> 
  as_tibble() |> 
  rename(x = V1, y = V2) |> 
   mutate(class = df_sak$class, name = df_sak$short_name) |> 
 ggplot(aes(x, y, colour = class |> as.factor(), label = class)) +
  geom_text(size = 2, family = "Fira Code") +
  labs(title = "UMAP continuous only",
      x = expression(paste("UMAP  ", italic(X))),
      y = expression(paste("UMAP  ", italic(Y))),
  ) +
  theme_pokemon +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )
```

# p4 pca cont fct

```{r}
p4 <- pca_sak_cont_fct$ind$coord |> 
  as_tibble() |> 
  rename(x = Dim.1, y = Dim.2) |> 
  mutate(class = df_sak$class, name = df_sak$short_name) |> 
  ggplot(aes(x, y, colour = class |> as.factor(), label = class)) +
  geom_text(size = 2, family = "Fira Code") +
  theme_pokemon +
  labs(
    x = paste0(
      "PCA 1, ",
      pca_sak_cont_fct$eig[1,2] |> round(digits = 1),
      "%"),
    y =  paste0(
      "PCA 2, ", 
      pca_sak_cont_fct$eig[2,2] |> round(digits = 1), 
      "%"
      ),
    title = "PCA continuous + factor 01"
  ) +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )

```

## p5: cont-fct tsne

```{r}
p5 <- tsne_sak_cont_fct |> 
  as_tibble() |> 
  rename(x = V1, y = V2) |> 
  mutate(class = df_sak$class, name = df_sak$short_name) |> 
  ggplot(aes(x, y, colour = class |> as.factor(), label = class)) +
  geom_text(size = 2, family = "Fira Code") +
  labs(
    title = "t-SNE continuous + factor 01",
    x = expression(paste("t-SNE  ", italic(X))),
    y = expression(paste("t-SNE  ", italic(Y))),
  ) +
  theme_pokemon +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )
```
## p6: umap
```{r}
p6 <- umap_sak_cont_fct |> 
  as_tibble() |> 
  rename(x = V1, y = V2) |> 
  mutate(class = df_sak$class, name = df_sak$short_name) |> 
  ggplot(aes(x, y, colour = class |> as.factor(), label = class)) +
  geom_text(size = 2, family = "Fira Code") +
  labs(
    title = "UMAP continuous + factor 01",
    x = expression(paste("UMAP  ", italic(X))),
    y = expression(paste("UMAP  ", italic(Y))),
  ) +
  theme_pokemon +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )
```
## p7: fct MCA
```{r}
p7 <- mca_sak_fct$ind$coord |> 
  as_tibble() |>
  rename(x = `Dim 1`, y = `Dim 2`) |> # slightly different from PCA() :(
  mutate(class = df_sak$class, name = df_sak$short_name) |> 
  ggplot(aes(x, y, colour = class |> as.factor(), label = class)) +
  geom_text(size = 2, family = "Fira Code") +
  theme_pokemon +
  labs(
    x = paste0(
      "MCA 1, ",
      mca_sak_fct$eig[1,2] |> round(digits = 1),
      "%"),
    y =  paste0(
      "MCA 2, ", 
      mca_sak_fct$eig[2,2] |> round(digits = 1), 
      "%"
      ),
    title = "MCA factor 01"
  ) +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )
```
## p8: umap fct

```{r}
p8 <- tsne_sak_fct |> 
  as_tibble() |> 
  rename(x = V1, y = V2) |> 
  mutate(class = df_sak$class, name = df_sak$short_name) |> 
  ggplot(aes(x, y, colour = class |> as.factor(), label = class)) +
  geom_text(size = 2, family = "Fira Code") +
  labs(
    title = "t-SNE factor 01",
    x = expression(paste("t-SNE ", italic(X))),
    y = expression(paste("t-SNE ", italic(Y))),
  ) +
  theme_pokemon +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )
```

## p9: umap fct 01

```{r}
p9 <- umap_sak_fct |> 
  as_tibble() |> 
  rename(x = V1, y = V2) |> 
  mutate(class = df_sak$class, name = df_sak$short_name) |> 
  ggplot(aes(x, y, colour = class |> as.factor(), label = class)) +
  geom_text(size = 2, family = "Fira Code") +
  labs(
    title = "UMAP factor 01",
    x = expression(paste("UMAP ", italic(X))),
    y = expression(paste("UMAP ", italic(Y))),
  ) +
  theme_pokemon +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )
```

## p10: fct MCA
```{r}
p10 <- 
  mca_sak_fct_cont$ind$coord |> 
  as_tibble() |>
  rename(x = `Dim 1`, y = `Dim 2`) |> # slightly different from PCA() :(
  mutate(class = df_sak$class, name = df_sak$short_name) |> 
  ggplot(aes(x, y, colour = class |> as.factor(), label = class)) +
  geom_text(size = 2, family = "Fira Code") +
  theme_pokemon +
  labs(
    x = paste0(
      "MCA 1, ",
      mca_sak_fct_cont$eig[1,2] |> round(digits = 1),
      "%"),
    y =  paste0(
      "MCA 2, ", 
      mca_sak_fct_cont$eig[2,2] |> round(digits = 1), 
      "%"
      ),
    title = "MCA factor 01 + binned cont"
  ) +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )
```
## p11: umap fct

```{r}
p11 <- 
  tsne_sak_fct_cont |> 
  as_tibble() |> 
  rename(x = V1, y = V2) |> 
  mutate(class = df_sak$class, name = df_sak$short_name) |> 
  ggplot(aes(x, y, colour = class |> as.factor(), label = class)) +
  geom_text(size = 2, family = "Fira Code") +
  labs(
    title = "t-SNE factor 01 + binned cont",
    x = expression(paste("t-SNE ", italic(X))),
    y = expression(paste("t-SNE ", italic(Y))),
  ) +
  theme_pokemon +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )
```

## p12: umap fct 01

```{r}
p12 <- 
  umap_sak_fct_cont |> 
  as_tibble() |> 
  rename(x = V1, y = V2) |> 
  mutate(class = df_sak$class, name = df_sak$short_name) |> 
  ggplot(aes(x, y, colour = class |> as.factor(), label = class)) +
  geom_text(size = 2, family = "Fira Code") +
  labs(
    title = "UMAP factor 01",
    x = expression(paste("UMAP ", italic(X))),
    y = expression(paste("UMAP ", italic(Y))),
  ) +
  theme_pokemon +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )
```
## p13: t-SNE, fct 01+cont as cont

```{r}
p13 <- 
  tsne_sak_fct_cont2 |> 
  as_tibble() |> 
  rename(x = V1, y = V2) |> 
  mutate(class = df_sak$class, name = df_sak$short_name) |> 
  ggplot(aes(x, y, colour = class |> as.factor(), label = class)) +
  geom_text(size = 2, family = "Fira Code") +
  labs(
    title = "t-SNE factor 01 + binned cont, perplexity = 5",
    x = expression(paste("t-SNE ", italic(X))),
    y = expression(paste("t-SNE ", italic(Y))),
  ) +
  theme_pokemon +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )
```
## p14: umap fct 01 +cont

```{r}
p14 <- 
  umap_sak_fct_cont2 |> 
  as_tibble() |> 
  rename(x = V1, y = V2) |> 
  mutate(class = df_sak$class, name = df_sak$short_name) |> 
  ggplot(aes(x, y, colour = class |> as.factor(), label = class)) +
  geom_text(size = 2, family = "Fira Code") +
  labs(
    title = "UMAP factor 01",
    x = expression(paste("UMAP ", italic(X))),
    y = expression(paste("UMAP ", italic(Y))),
  ) +
  theme_pokemon +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )
```

## p14a: umap factorize

test if converting dbl 0L/1L to factor changes something. doubt it

... it does, drastically.

```{r}
p14a <- 
  umap_sak_fct_cont2_test_factorize |> 
  as_tibble() |> 
  rename(x = V1, y = V2) |> 
  mutate(class = df_sak$class, name = df_sak$short_name) |> 
  ggplot(aes(x, y, colour = class |> as.factor(), label = class)) +
  geom_text(size = 2, family = "Fira Code") +
  labs(
    title = "UMAP factor 01",
    x = expression(paste("UMAP ", italic(X))),
    y = expression(paste("UMAP ", italic(Y))),
  ) +
  theme_pokemon +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )
```


## umap metric: categorical + euclidean,

でもcategoricalなデータは01をファクター化しただけのもの（wide)。

```{r}
umap_sak_fct_cont2_modified_metric |> 
  as_tibble() |> 
  rename(x = V1, y = V2) |> 
  mutate(class = df_sak$class, name = df_sak$short_name) |> 
  ggplot(aes(x, y, colour = class |> as.factor(), label = class)) +
  geom_text(size = 2, family = "Fira Code") +
  labs(
    title = "UMAP factor 01",
    x = expression(paste("UMAP ", italic(X))),
    y = expression(paste("UMAP ", italic(Y))),
  ) +
  theme_pokemon +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )
```

## p15: umap, rich factor + factor(0/1 factorized) + continuous

```{r}
p15 <- umap_fct_cont3 |> 
  as_tibble() |> 
  rename(x = V1, y = V2) |> 
  mutate(class = df_sak$class, name = df_sak$short_name) |> 
  ggplot(aes(x, y, colour = class |> as.factor(), label = class)) +
  geom_text(size = 2, family = "Fira Code") +
  labs(
    title = "UMAP richest",
    x = expression(paste("UMAP ", italic(X))),
    y = expression(paste("UMAP ", italic(Y))),
  ) +
  theme_pokemon +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )
```
