---
title: "R Notebook"
output: html_notebook
---
# narrative & grid search

やりたい
PCA Cont only
PCA converted all to cont
Correspondence Fct only
correspondence converted to factors (forced dichotomization)
t-SNE mixed (possible or not)
t-SNE cont only
UMAP fct/euclidean mixed (uwot)
UMAP converted to cont, euclidean

x

datasets: pokémon; swiss? janiota?


# packages

{FactoMineR} for PCA,
{FactoMineR} for Corresp,
{tsne} for t-SNE,
{uwot} for UMAP

# library

```{r}
library(tidyverse)
library(googlesheets4)
library(FactoMineR)
library(tsne)
library(uwot)
library(patchwork)
library(paletteer)
```
# predefined funcs
```{r}
theme_pokemon <- 
  theme_minimal(base_size = 6, base_family = "Helvetica") +
  theme(
    line = element_line(size = .15/.751, colour = "black"),
    text = element_text(size = 6, colour = "black"),
    title = element_text(size = 6, colour = "black"),
    panel.grid = element_blank(),
    axis.line = element_line(),
    axis.text = element_text(size = 6, colour = "black"),
    axis.ticks = element_line(),
    plot.background = element_rect(fill = "white", colour = NA),
    strip.text = element_text(size = 6),
    legend.text = element_text(size = 6),
    legend.key.size = unit(3, units = "mm")
  )
```


# datasets


```{r}
data_sak_raw <- read_sheet("https://docs.google.com/spreadsheets/d/1CMR07q3v-0UwosNtEHKAyekOpQTRj7dpLUb323dFt-E/edit#gid=0", sheet = "swiss army knives")
# df_pokemon <- read_sheet("https://docs.google.com/spreadsheets/d/155e3yO6J-l4tCIF5J0cwv2tWaUTzq5RPkMJZOiA0dRI/edit#gid=0", sheet = cards) # equivalent to df_gene in pokemon tcg

```

toriaezu pokemon de hitotoori yatte tarinakattara hoka yaru?-
->too few continuous variables
```{r}

df_for_umap <- df_gene |> 
  filter(!is.na(pokedex_id)) |> 
  mutate(
    seconds_since_first_release =
      interval(min(release_date), release_date) |>
      int_length(),
    is_secret_card = replace_na(is_secret_card, FALSE),
    time_since_first_release_standardized = seconds_since_first_release / seconds_since_first_release |> max(),
    master = replace_na(master, "no_master"),
    forme_variant = replace_na(forme_variant, "no_forme_variant")
  ) |> 
  select(
    release_date, # for plot,
    time_since_first_release_standardized, pokemon_gen, # continuous values
    pokedex_id, card_type2, series, is_secret_card, series_class, is_gx:forme_variant, # factors
    ) |>
  mutate(
    across(pokemon_gen:pokemon_gen, ~as.numeric(.x)), # continuous values
    across(pokedex_id:forme_variant, ~as.factor(.x)), # factors
  )
df_for_umap # 12724 x 43

```



## run UMAP


### prepare 

[uwot](https://github.com/jlmelville/uwot) mate
https://www.rdocumentation.org/packages/uwot/versions/0.1.11/topics/umap
{uwot}'s `uwot::umap()` does not accept tibble. convert it to data.frame
```{r}
set.seed(1121)
df_for_umap2 <- df_for_umap |> 
  select(-release_date)
columns <- df_for_umap2 |> colnames()
```

### run 

```{r fig.height = 5, fig.width = 5}
matrix_umap <- df_for_umap2 |> 
  as.data.frame() |> 
  uwot::umap(
    verbose = TRUE, 
    n_threads = 6,
    n_neighbors = 15, # recommended: 2-100 default 15. larger = slower
    min_dist = .9, # set 0-1, set smaller than spread. usually control this instead of spread
    spread = 40, # default 1
    fast_sgd = TRUE, # TRUE for viz 
    metric = list("euclidean" = columns[1:2],
                  "categorical" = columns[3:42])
  )
# prepare df for plot
df_umap <- matrix_umap |> 
  as_tibble() |> 
  rename(x = V1, y = V2) |> 
  rowid_to_column() |> 
  left_join(df_for_umap |> rowid_to_column(), by = "rowid")

### plot
 df_umap |> 
  ggplot(
    aes(
      x = x,
      y = y, 
      fill = release_date
    )
  ) +
  geom_point(pch = 21, colour = "black", size = .5, stroke = 0) +
  scale_x_continuous(
    breaks = c(-100, 0, 100),
    labels = c(-100, "", 100),
    name = expression(paste("UMAP  ", italic(X)))
  ) +
  scale_y_continuous(
    breaks = c(-100, 0, 100),
    labels = c(-100,  0, 100),
    name = expression(paste("UMAP  ", italic(Y)))
  ) +
  # scale_fill_paletteer_c(
  #   "pals::ocean.deep",
  # ) +
  scale_fill_viridis_c() +
  facet_grid(
    cols = vars(card_type2),
    rows = vars(pokemon_gen)) +
  theme_pokemon +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )
```
